---
title: "8-  Hierarchical Clustering"
author: "Allex Belostrino"
date: "11/25/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read the data

```{r}
#install.packages("tidyverse")
#install.packages("readr")
Cereals <- read.csv("/Users/allexnicolebelostrino/Desktop/Assignment_5/Cereals.csv")

#Remove cereals with missing values
Cereals <- na.omit(Cereals)
head(Cereals)
summary(Cereals)
```

After cleaning the data, we see that there were three cereals with missing values. Therefore, giving us a remaining 74 cereals from the original 77.

Use dataset only with numerical variables and standardize

```{r}
numeric_cereal <- Cereals[,c(4:16)]
scaled_numeric_cereal <- scale(numeric_cereal)

```

Compute Euclidean distance

```{r}
d <- dist(Cereals, method = "euclidean")

```

```{r}
# normalize input variables
Cereals.norm <- sapply(Cereals[,c(4:16)], scale)

# add row names: cereals
row.names(Cereals.norm) <- row.names(Cereals) 


```

```{r}
#install.packages(stats)
#compute normalized distance based on all variables
Cereals.norm <- dist(Cereals.norm, method = "euclidean")

# in hclust() set argument method =  
# to "ward.D", "single", "complete", "average"
hc1 <- hclust(Cereals.norm, method = "single")
plot(hc1, hang = -1, ann = FALSE)
hc2 <- hclust(Cereals.norm, method = "average")
plot(hc2, hang = -1, ann = FALSE)
hc3 <- hclust(Cereals.norm, method = "complete")
plot(hc3, hang = -1, ann =FALSE)
hc4 <- hclust(Cereals.norm, method = "ward.D")
plot(hc4, hang = -1, ann= FALSE)
```

Use AGNES to compare the clusters

```{r}
#install.packages(cluster)
library(cluster)

#Compute with AGNES and different linkage methods
hc_single <- agnes(Cereals, method ="single")
hc_average <- agnes(Cereals, method ="average")
hc_complete <- agnes(Cereals, method ="complete")
hc_ward <- agnes(Cereals, method = "ward")

#Compare agglomerative coefficients
print(hc_single$ac)
print(hc_average$ac)
print(hc_complete$ac)
print(hc_ward$ac)


pltree(hc_single, cex = 0.6, hang =-1 , main = "Dendrogram of Agnes")
pltree(hc_average, cex = 0.6, hang =-1 , main = "Dendrogram of Agnes")
pltree(hc_complete, cex = 0.6, hang =-1 , main = "Dendrogram of Agnes")
pltree(hc_ward, cex = 0.6, hang =-1 , main = "Dendrogram of Agnes")
```

The best method is the Ward method as is the strongest clustering structure.

Cutting dendrograms 
```{r}
#single linkage. Height is based off of 80. Therefore, we will choose to have 2 clusters
memb <- cutree(hc1, k = 2)
memb
```

```{r}
#average linkage. Height is based off of 150. Therefore we will choose to have 4 clusters
memb <- cutree(hc2, k = 4)
memb

```

```{r}
#complete linkage. Height is based off of 200. Therefore, we will choose to have 5 clusters
memb <- cutree(hc3, k = 5)
memb
```

```{r}
#ward. Height is based off of 600. Therefore, we will choose to have 3 clusters.
memb <- cutree(hc4, k = 7)
memb
```


Comment on the structure of the clusters and on their stability
```{r}
#install.packages("caret")
library(caret)
#Combine cluster from question 1
NewCereal <- rbind(hc1,hc2,hc3,hc4)

#Making partitions
set.seed(123)
PartA <- sample(nrow(Cereals.norm), replace = TRUE)
PartB <- 3 - PartA

#Fitting PartA cluster
clusterA <- cutree(hc_ward, k = 7)

#Assign PartA centroids to PartB
clusterB <- cutree(hc_ward, newdata = Cereals.norm[PartB, ] )

#Assess consistency
consistency <- sum(PartA == clusterB) / length(clusterB)
```



The elementary public schools would like to choose a set of cereals to include in their daily cafeterias. Find a cluster of "healthy cereals." Should the data be normalized? Which cluster is doing great on healthy cereals?

```{r}

#list of characteristics
healthy.cereals <- NewCereal[which(NewCereal$calories < 50 & NewCereal$protein > 4 & NewCereal$fat < 2 & NewCereal$fiber > 3 & NewCereal$carbo < 15 & NewCereal$sugar < 8 & NewCereal$potass > 100 & NewCereal$vitamins > 25)]


```
Normalization is needed since there is a range of characteristics needed to determine healthy cereals.