---
title: "Final Exam by Allex Nicole Belostrino"
output: html_document
date: "2023-12-09"
---

Import Fuel Deliveries receipts and costs dataset. But first, remove columns with significant missing values. In this case it is mercury_content_ppm (Column Q), secondary_transportation_mode_code(Column R), and natural_gas_delivery_contract_type_code(Column T). Initially we have 21 columns, including the rowiD. This time, after removing columns, we will expect to have 18 columns total including rowID.
```{r}
Power <- read.csv("/Users/allexnicolebelostrino/Desktop/Final Exam Folder/Power.csv")
Power2 <- Power[ , -c(15,18,20)]
ncol(Power2)
nrow(Power2)

```


Let's ensure our dataset has both the right attributes like numerical, character, and integer.
```{r}
str(Power2)

```


Randomly sample 2% of the dataset 
```{r}
set.seed(3548)
total_rows <- nrow(Power2)
sample.size <- round(0.02 * total_rows)
sample.data <- Power2[sample(1:total_rows, size = sample.size, replace = FALSE), ]

```


Partition 70% of the sample data into training and the rest as test.
```{r}
#install.packages("caret")
#install.packages("ggplot2")
#install.packages("lattice")
library(caret)

#Create index for data partitioning
set.seed(3548)
index <- createDataPartition(y = 1:nrow(sample.data), p = 0.7, list = FALSE)

#Create training and testing
train <- sample.data[index, ]
test <- sample.data[-index, ]
```


Training set should be about 9,000 and test set should be about 3,000
```{r}

#Training set
length(train)


#Test Set
length(test)

nrow(train)
nrow(test)
```
When checking from the data frame, training has 9,291 items while test has 3,980. Both sets have 18 columns.


Question 1: Which type of transportation is mostly used for delivering different fuels? (Coal, Gas, and Oil)
```{r}
transpo.data <- table(sample.data$fuel_type_code_pudl, sample.data$primary_transportation_mode_code)
barplot(transpo.data, beside = TRUE,
        legend.text = TRUE,
        col = c("pink", "skyblue", "yellow"),
        main = "Which type of transportation is used for delivering different fuels?",
        xlab = "Transportation Modes" , ylab = "Frequency")

```
In this bar plot, we compare the three types of fuels and their transportation differences. 


Question 2: What information is revealed by clustering? Use any and all variables to describe and identify the clusters. 
```{r}

#grab data
set.seed(3000)
sample.data <- data.frame(
  ash_content_pct = rnorm(1000),
  fuel_received_units = rnorm(1000)
) 

#choose number of k
k <- 4

#do the clustering
kmeans.result <- kmeans(sample.data, centers = k)

cluster.assign <- kmeans.result$cluster

centroids <- kmeans.result$centers


#do the viz for kmeans
plot(
  sample.data$ash_content_pct,
 sample.data$fuel_received_units,
  col = cluster.assign,
  pch = 20,
  main = "Cluster of fuels received by plant and its ash content",
  xlab = "Ash Content",
  ylab = "Fuel Received in units"
)

points(centroids[ , 1], centroids[ ,2], col = 1:k, pch = 3, cex = 2)

```
for this cluster, I chose k = 4. However, by mere looking at the plot, I can tell that I should retain or lower the number of k since there are points that seem to be close to one another that should be a part of a cluster that is too far from them. If we chose a smaller number, we would have a more sensitive and accurate grouping.


Question 3: How should the value for the number of clusters be chosen?
```{r}

#Elbow method
#install.packages("factoextra")
library(factoextra)
sample.data <- scale(sample.data)
fviz_nbclust(sample.data, kmeans, method = "wss")

```
For choosing the optimal number of k, it is evident that in the elbow method there is a bend at k = 3


```{r}

#Silhouette method
fviz_nbclust(sample.data, kmeans, method = "silhouette")


```
For this method, it suggests that the optimal number of k should be 7, which was the same as our theory in the previous chunk.


We will then re-do the clustering using k=3
```{r}

set.seed(3000)
sample.data <- data.frame(
  ash_content_pct = rnorm(1000),
  fuel_received_units = rnorm(1000)
) 

#choose number of k
k <- 3

#do the clustering
kmeans.result <- kmeans(sample.data, centers = k)

cluster.assign <- kmeans.result$cluster

centroids <- kmeans.result$centers


#do the viz for kmeans
plot(
  sample.data$ash_content_pct,
 sample.data$fuel_received_units,
  col = cluster.assign,
  pch = 20,
  main = "Cluster of fuels received by plant and its ash content",
  xlab = "Ash Content",
  ylab = "Fuel Received in units"
)

points(centroids[ , 1], centroids[ ,2], col = 1:k, pch = 3, cex = 2)
```